---
title: "EDLD 610 Winter 2020 Lab 3"
author: "Jim Wright and Claire Zhang"
date: "1/29/2020"
output: html_document
toc: true
toc_float: true 
theme: united 
highlight: zenburn 
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      fig.width = 10,
                      fig.height = 10,
                      dev.args = list(bg="transparent"))

library(tidyverse)
theme_set(theme_minimal())


library(fivethirtyeight)
```


# Prompt 1: Google Trends Data
```{r prompt 1, include=TRUE}
head(google_trends)

ggplot(google_trends, aes(date, hurricane_harvey_us)) +
  geom_area(fill = "cornflowerblue", 
            alpha = 0.3) +
  geom_line(lwd = 1.6, 
            color = "gray40") +
    geom_smooth(color = "magenta",
              lwd = 1.4,
              se = FALSE) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Change in Trend",
       title = "Hurricane Harvey")

g_trend <- google_trends %>% 
  pivot_longer(
    cols = c(2:5),
    names_to = "hurricane",
    values_to = "trends"
  )

ggplot(g_trend, aes(date, trends)) +
 geom_line(aes(color = hurricane), 
           lwd = 1.6) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Change in Trend",
       title = "Hurricanes Under Trump Administration") 
  
ggplot(g_trend, aes(date, trends)) +
 geom_area(fill = "cornflowerblue", 
            alpha = 0.9) +
 geom_line(aes(color = hurricane), 
           lwd = 1.6) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Change in Trend",
       title = "Google Trends of Hurricanes Under the Trump Administration") 

ggplot(g_trend, aes(date, trends)) +
  geom_ribbon(aes(fill = hurricane, ymin = 0, ymax = trends),
              alpha = 0.7) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Change in Trend",
       title = "Google Trends of Hurricanes Under the Trump Administration") 

ggplot(g_trend, aes(date, hurricane)) +
  geom_tile(aes(fill = trends)) +
  scale_fill_viridis_c(option = "inferno") +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Hurricane",
       title = "Google Trends of Hurricanes Under the Trump Administration") 



landfall <- tibble(
  date = lubridate::mdy(c(
    "August 25, 2017",
    "September 10, 2017",
    "September 20, 2017"
  )),
  hurricane = c("Hurricane Landfall",
                "Irma Landfall",
                "Maria Landfall")
)

head(landfall)

ggplot(g_trend, aes(date, trends)) +
  geom_ribbon(aes(fill = hurricane, ymin = 0, ymax = trends),
              alpha = 0.7) +
    geom_vline(aes(xintercept = date), landfall,
             color = "black", 
             linetype = "longdash") +
   geom_text(aes(x = date, y = 50, label = hurricane), landfall,
            color = "black") +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Change in Trend",
       title = "Google Trends of Hurricanes Under the Trump Administration") 



```


# Prompt 2: Google Trends Data
```{r prompt 2, include=TRUE}
head(tv_states)

tv <- tv_states %>% 
   pivot_longer(
    cols = c(2:4),
    names_to = "state",
    values_to = "percentage"
  )
  
head(tv)

ggplot(tv, aes(date, percentage)) +
  geom_ribbon(aes(fill = state, ymin = 0, ymax = percentage), color = "white") +
  scale_fill_manual(values = c("orange","magenta", "blue")) +
  geom_vline(aes(xintercept = date), landfall,
             color = "gray50", 
             linetype = "longdash",
             alpha = 0.2) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Date",
       y = "Share of sentences",
       title = "National cable news networks") 
  

```

# Comic-Characters Data Set
```{r prompt 3, include=TRUE}
library(lubridate)
library(janitor)

# Get DC characters:
comic_characters_dc <- 
  "https://github.com/fivethirtyeight/data/raw/master/comic-characters/dc-wikia-data.csv" %>% 
  read_csv() %>% 
  clean_names() %>% 
  mutate(publisher = "DC")

# Get Marvel characters:
comic_characters_marvel <- 
  "https://github.com/fivethirtyeight/data/raw/master/comic-characters/marvel-wikia-data.csv" %>% 
  read_csv() %>% 
  clean_names() %>% 
  mutate(publisher = "Marvel")

# Merge two dataset and perform further data wrangling:
comic_characters <-
  comic_characters_dc %>% 
  bind_rows(comic_characters_marvel) %>% 
  separate(first_appearance, c("year2", "month"), ", ", remove = FALSE) %>%
  mutate(
    # If month was missing, set as January and day as 01:
    month = ifelse(is.na(month), "01", month),
    day = "01",
    # Note some years missing:
    date = ymd(paste(year, month, day, sep = "-")),
    align = factor(
      align, 
      levels = c("Bad Characters", "Reformed Criminals", "Netural Characters", "Good Characters"),
      ordered = TRUE)
  ) %>%
  select(publisher, everything(), -c(year2, day))

head(comic_characters)
comic_characters$sex

comic_characters %>% 
  count(sex)

ggplot(comic_characters, aes(sex)) +
  geom_bar(fill = "blue", width = 1) +
  theme_minimal() +
  coord_flip() +
  scale_x_discrete(limits = rev(levels(comic_characters$sex))) +
  theme(plot.title = element_text(color = "black", size = 12, face = "bold", hjust = 0.5)) +
  labs(x = "Total",
       y = "Character Gender",
       title = "Total Number Comic Characters by Gender") 

percentages <- comic_characters %>% 
  count(sex) %>% 
  mutate(percent = n/sum(n)*100)

head(percentages)

ggplot(percentages, aes(x = 10, y = percent)) +
  geom_col(aes(fill = sex)) +
  coord_flip()

ggplot(percentages, aes(sex, percent)) +
  geom_col(fill = "blue", width = 1) +
  coord_flip()

ggplot(percentages, aes(sex)) +
  geom_bar(aes(y = ..prop.., fill = factor(sex)), stat = "count") +
  coord_flip() +
  geom_text(aes(label = scales::percent(..prop..),
                   y= ..prop.. ), stat= "count", vjust = -.5) +
      scale_y_continuous(labels = scales::percent) +
  guides(fill = "none") +
  labs(title = "Comic Book Character Gender",
       y = "Percent")
```

Hi Jim,
Love your ideas! Great work! I learned a lot from your code.
Here's something I did today.Hopefully I'll have time to do more over the weekend.
Claie

1. Start with the google trends data.

Visualize the change in trends for each hurricane over time in one plot with three scales. Map color to a discrete scale.

Visualize the trends again, this time with solid shapes (if your first plot was already solid shapes, use lines). Hint, geom_area can be difficult to work with if you have a grouping factor. Try geom_ribbon instead.

Visualize the same relation, but change the scales so color is mapped to a continuous scale.

Visualize the change in trends for each hurricane over time in one plot with three scales. Map color to a discrete scale.

```{r}
a <- google_trends %>%
  pivot_longer(2:4,
               names_to = "hurricane",
               values_to = "trends")

ggplot(a, aes(date,trends))+
  geom_line(aes(color = hurricane))

ggplot(a, aes(date, trends))+
  geom_ribbon(aes(fill = hurricane, ymin = 0, ymax = trends),
              alpha = 0.9,
              color = "grey70")

ggplot(a, aes(date, hurricane)) +
  geom_tile(aes(fill = trends),
            color = "gray70") +
  scale_fill_viridis_c(option = "magma") +
  coord_fixed()+
  coord_flip()+
  theme(legend.position = "right",
        legend.direction = "vertical",
        legend.key.height = unit(.5,"cm"))
```

Create a data frame that includes the dates that each hurricane made landfall.

Annotate one of your first two plots to include these reference points with vertical lines. You do not have to include text annotations - just the vertical lines. But if you would like to try, look into geom_text or geom_label. The hurricanes made landfall on (approximately) the following days:
  
  Harvey: August 25, 2017
Irma: September 10, 2017
Jose: Did not make landfall
Maria: September 20, 2017

Make any final modifications to the plot you’d like to communicate the information as effectively as possible.

```{r}
landfall <- tibble(date = lubridate::mdy(c("August 25, 2017", 
                                           "September 10, 2017", 
                                           "September 20, 2017")),
                   hurricane = c("Harvey Landfall", 
                                 "Irma Landfall", 
                                 "Maria Landfall"))

ggplot(a, aes(date, trends))+
  geom_ribbon(aes(fill = hurricane, ymin = 0, ymax = trends),
              alpha = 0.9,
              color = "grey30")+
  geom_vline(aes(xintercept = date), landfall,
             color = "#fb6107", 
             lty = "dashed") +
  geom_text(aes(x = date, y = 80, label = hurricane), landfall,
            color = "grey40",
            nudge_x = 2,
            size = 3,
            angle = 45)+
  labs(x = "Date", 
       y = "Google Trends",
       title = "Hurricane Google trends over time")

```

2. Replicate the “National cable news networks” plot from the story using the tv_states data. Don’t worry about all the labels. Try to match the colors but don’t worry about getting them exact.

```{r}
t <- tv_states %>% 
  gather(state, percentage, -date) %>% 
  mutate(state = factor(state, levels = c("florida", "texas", "puerto_rico"))) 

ggplot(b, aes(date, percentage)) +
  geom_ribbon(aes(fill = state, ymin = 0, ymax = percentage), color = "grey70")+
  geom_vline(aes(xintercept = date), landfall,
             color = "#fb6107", 
             lty = "dashed")+
  geom_text(aes(x = date, y = 3.2, label = hurricane), landfall,
            color = "grey40",
            nudge_x = 2,
            size = 3,
            angle = 45)

```

3. Use the comic_characters dataset to visualize the proportion of different genders using at least three different methods. .

```{r}
# Note that the {fivethirtyeight} package only includes a preview (10 rows) of the data. Run the following code to get access to the full dataset.
# Get DC characters:
comic_characters_dc <- 
  "https://github.com/fivethirtyeight/data/raw/master/comic-characters/dc-wikia-data.csv" %>% 
  read_csv() %>% 
  clean_names() %>% 
  mutate(publisher = "DC")

# Get Marvel characters:
comic_characters_marvel <- 
  "https://github.com/fivethirtyeight/data/raw/master/comic-characters/marvel-wikia-data.csv" %>% 
  read_csv() %>% 
  clean_names() %>% 
  mutate(publisher = "Marvel")

# Merge two dataset and perform further data wrangling:
comic_characters <-
  comic_characters_dc %>% 
  bind_rows(comic_characters_marvel) %>% 
  separate(first_appearance, c("year2", "month"), ", ", remove = FALSE) %>%
  mutate(
    # If month was missing, set as January and day as 01:
    month = ifelse(is.na(month), "01", month),
    day = "01",
    # Note some years missing:
    date = ymd(paste(year, month, day, sep = "-")),
    align = factor(
      align, 
      levels = c("Bad Characters", "Reformed Criminals", "Netural Characters", "Good Characters"),
      ordered = TRUE)) %>%
  select(publisher, everything(), -c(year2, day))
```


With your final visualization, make any additional modifications you think would help reduce the cognitive load necessary to comprehend the figure

```{r}
percents <- comic_characters %>% 
  count(sex) %>% 
  mutate(percent = n/sum(n)*100)

c <- na.omit(left_join(comic_characters, percents))

ggplot(c, aes(x = fct_reorder(sex, n), y = percent,)) +
  geom_col(aes(fill = fct_reorder(sex, n)), position = "dodge") +
  labs(x = "Gender",
       y = "Percentage")+
  theme_minimal()+
  coord_flip()
```
